### 2016-04-08 ms/ms_probaweb.tex
Spencer edits

% Table 1: example for a pair of species
% Table 2: model comparison across all species

@@

\textbf{Title:} Bringing Elton and Grinnell together: a quantitative framework to represent the biogeography of ecological interactions

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Ben Baiser$^{3}$, Jennifer A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo
Martinez$^{6}$, Tommy Nyman$^{7}$, Timoth\'ee Poisot$^{2,8}$,  Spencer Wood$^{9}$, Daniel B. Stouffer$^{10}$, Jason Tylianakis$^{10,11}$ Tomas Roslin$^{12}$,\\

@@

9:\\

@@

Biogeography has historically focused on the spatial distribution and
abundance of species, neglecting variation in the way species interact with
each other. Models of species distribution and of interactions have
historically taken different paths. Here, we plea for an integrated approach,
adopting the view that community structure is best represented as an
ecological network of interactions. We outline conceptual approach, suggesting
that the ecological niche may be redefined to encompass the effect of the
environment on species distribution (the Grinnellian dimension of the niche)
and on the ecological interactions among them (the Eltonian dimension).
Starting from this novel concept, we develop a quantitative theory to explain
turnover of interactions in space and time – i.e. a novel approach to
interaction distribution modelling. We decompose the probability with which
two species interact with each other in two parts: the probability with which
the species co-occur at a given location, and the probability that if they do
so, they will interact. Both of these probabilities could be conditional on
the environment. We apply this novel framework to a large data set of
host–parasite interactions across Europe and find that two aspects of the
environment (temperature and precipitation) leave a strong imprint on species
co-occurrence, but not on the probability of local interactions. Even where
species co-occur, interaction proves a stochastic rather than deterministic
process, adding to variation in realized network structure. We also find that
a large majority of species pairs are never found together, thus precluding
any inferences regarding their probability to interact. Our framework provides
a first conceptual framework to explain the variation of network structure at
large spatial scales and opens new perspectives at the frontier between
species distribution modelling and community ecology.

@@

Community ecology is defined in most textbooks as \textit{the study of the
interactions that determine the distribution and abundance of organisms}
\citep{Krebs2001}. Despite a general consensus on this definition
(Scheiner2007), it is surprising that research on variation in community
structure has focused mostly on the turnover of species composition
\citep{Anderson2011}, neglecting variation in the way species interact with
each other \citep{Poisot2015a}. Given this omission, it is perhaps not
surprising that biogeographers are still struggling to establish whether
interactions actually impact the distribution of species \citep{Wisz2012,
Kissling2012}. Recent attempts at accounting for interactions in species
distribution models \citep{Pollock2014, Pelissier2013} have brought some
methodological advances. Yet, these techniques are still based on a ’species-
based’ approach to communities, where interactions are merely treated as fixed
covariates affecting distribution.

As a more explicit description of interaction structure among species, the
network approach offers a convenient representation of communities. Species
are represented as nodes and interactions by links. To date, studies of
network diversity have mostly been concerned with the distribution of
interactions within locations, and less so with the variation among locations
\citep{Dunne2005, Bascompte2007, Ings2007, Kefi2012}. However, there is now
ample evidence that interaction networks vary in space and time
\citep{Poisot2012, Albouy2014, Poisot2016, Trojelsgaard2015}. Metacommunity
theory provides explanation for the variation in the distribution of the
different nodes \citep{Gravel2011c, Pillai2011}, but there is no explanation
to the joint variation of node and link occurrences. We urgently need a
conceptual framework to formalize these observations.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is an evident need to bring the two together.
Here, we plea for an integrated approach, adopting the view that community
structure is best represented as an ecological network of interactions. Based
on this idea, we propose a new description of the basic concept of the
ecological niche, now integrating the effect of the environment on species
distribution and on the ecological interactions among them. Starting from this
redefined concept, we develop a unified theory to explain turnover of
interactions in space and time. We first present the conceptual framework, and
then formalize it mathematically, using a probabilistic model to represent the
sampling of the regional pool of interactions. At the level of species pairs,
the statistical approach could be conceived as an interaction distribution
model. At the community level, the approach provides a likelihood-based method
to compare different hypotheses of network turnover. By applying this novel
framework to a large data set on host–parasite interactions across Europe, we
find that variation of the environment causes turnover of both species and
interactions. The network structure changes systematically across the
latitudinal gradient, with a peak of connectance at intermediate latitudes.

@@

The problem of community assembly is often formulated as \textit{how do we
sample a regional pool of species to constitute a local community?} This
question could be rewritten to address the problem of network assembly, as
\textit{how do we sample a regional pool of interactions to constitute a local
interaction network?} An illustration of this problem for a food web is
provided in Figure 1. The metaweb represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 275 nodes, and 1173 links among plants (52 nodes), herbivores (96 nodes),
and parasitoids (127 nodes) from Northern Europe. An instance of a local
community is also illustrated, with 45 nodes and 93 interactions. Only 55.0\%
of all potential interactions are realized. Our objective here is to provide a
conceptual framework to explain the sampling of the regional pool of
interactions, along with a quantitative method to predict it. The problem
could be formalized sequentially by first understanding why only a fraction of
the species are co- occurring locally, and then why these species interact or
not.

There are multiple causes of spatial turnover in community composition. The
first and most-studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among sites by selecting the locally
fittest species \citep{Leibold2004}. Stochasticity plays an additional role,
either through the inherently unpredictable nature of colonization and
extinction events (Hanski1999), or through strong non-linear feedbacks
generating alternative transients and equilibriums \citep{Chase2007,
Vellend2014}. Analyses of community turnover are usually performed with data
represented in a table with rows corresponding to sites (or measurements) and
columns to species. Metrics of beta diversity quantify the variance of this
community data \citep{Legendre2005}. Traditional approaches rely on measures
of dissimilarity among communities, such as the Jaccard or Bray–Curtis
indices. A more recent approach decomposes the total variation of the
community data into species and site contributions to beta diversity
\citep{Legendre2013}. Even though these methods compare whole lists of species
among sites or measurements, they remain fundamentally ’species-based’, since
they report the variation within columns. None of them explicitly considers
the variation of associations (i.e., of pairs or higher- order motifs –
\citealt{Stouffer2007}).

The “niche” is by far the dominant concept to explain species distributions
and community assembly, from the local to the global scale. Following
\citealt{Hutchinson1957}, the niche is viewed as the set of environmental
conditions allowing a population to establish and persist (see also
\citealt{Holt2009}. Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to operationalize with species distribution models, as it maps
naturally on available distributional and environmental data. In consequence,
a vast array of statistical tools have been developed to implement it (e.g.
Biomod \citealt{Thuiller2003}, MaxEnt \citealt{Phillips2006}). It is however
much harder to account for ecological interactions with this approach
(Peterson2011). As such, these interactions are often viewed as externalities
constraining or expanding the range of environmental conditions required for a
species to maintain a viable population \citep{Pulliam2000, Soberon2007}.

@@

Interestingly, the ecological network literature also has its own ’niche
model’ to position a species in a community (Williams2000). The niche of a
species in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over 3 to 5 different niche axes \citep{Williams2000, Eklof2013}.
The niche model of food web structure has successfully explained the
complexity of a variety of networks, from food webs to plant–pollinator
systems \citep{Allesina2008, Williams2010, Eklof2013}. This
conceptual framework is, however, limited to local communities, and does not
provide any explanation for the turnover of network structure along
environmental gradients.

@@

A more integrative description of the niche is key to understand spatial and
temporal turnover in community structure. Despite several attempts to update
the concept of the ecological niche, ecologists have not moved far past the
“n-dimensional hypervolume” defined by Hutchinson. Despite its intuitive
interpretation and easy translation into species distribution models
\citep{Boulangeat2012, Blonder2014}, the concept has been constantly
criticized \citep{Hardin1960, Peters1991, Silvertown2004}, and several
attempts have been made to expand and reinforce it \citep{Pulliam2000,
Chase2003, Soberon2007, Holt2009, McInerny2012b}.

Part of the problem surrounding the niche concept has been clarified with the
distinction between Eltonian and Grinnellian definitions (Chase2003). The
Grinnellian dimension of the niche is the set of environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is intuitive to apply, and constitutes the conceptual backbone of
species distribution models. The Eltonian niche, on the other hand, is the
effect of a species on its environment. This aspect of the niche is well known
by community ecologists, but is trickier to turn into predictive models.
Nonetheless, the development of the niche model of food web structure
\citep{Williams2000} and its parameterization \citep{Williams2010; Gravel2013}
made it more operational.

These perspectives are rather orthogonal to each other, which has resulted in
considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile them in their definition of the
niche: “[The niche is] the joint description of the environmental conditions
that allow a species to satisfy its minimum requirements so that the birth
rate of a local population is equal to or greater than its death rate along
with the set of per capita effects of that species on these environmental
conditions”. Their representation merges zero-net growth isoclines delimiting
the Grinnellian niche (when do populations persist) with impact vectors
delimiting the Eltonian niche (what is the per-capita impact). While this
representation has been very influential in local-scale community ecology (the
resource-ratio theory of coexistence, \citealt{Tilman1982}, it remains
impractical at larger spatial scales because of the difficulties to measure
it. The absence of any mathematical representation of the niche that could
easily be fitted to ecological data may explain why biogeographers are still
struggling to develop species distribution models that also consider
ecological interactions.

We propose to integrate the two perspectives of the niche with a visual
representation of both components. The underlying rationale is that, in
addition to the environmental constraints on demographic performance, any
organism requires resources to sustain its metabolic demands and reproduction.
Abiotic environmental axes are any non-consumable factors affecting the
demographic performance of an organism. Alternatively, the resource axes are
traits of the resources allowing interactions with the consumer. The niche
should therefore be viewed as the set of abiotic environmental factors (the
Grinnellian component) along with the set of traits (the Eltonian component)
allowing a population to establish and to persist at a location. Accordingly,
each species can be characterized by an optimal position in both the
environmental (x-axis) and the trait (y-axis) plane. The integrated niche is
then the hypervolume where interactions can occur and sustain a population.
This approach radically changes the representation of the niche, putting the
distributions and ecological interactions of species in the same formalism.
The limits of the niche axes could be independent of each other (as in the
example of Fig. 2), but they could also interact. For instance, the optimal
prey size for predatory fishes could decrease with increasing temperature
\citep{Lelong2015}, which would make diet boundaries functions of the
environment. The other way around, we could also consider that the growth rate
of the predator changes with the size of its prey items, thereby altering the
environmental boundaries.

%========================================================%
\section*{A probabilistic representation of ecological interactions networks in space}

We now formalize the integrated niche with a probabilistic approach to
interactions and distributions. We seek to represent the probability that an
interaction between species $i$ and $j$ occurs at location $y$. We define $L_{ijy}$ as a
stochastic variable, and are looking at the probability that this event
occurs, $P(L_{ijy})$. The occurrence of an interaction is dependent on the co-
occurrence of species $i$ and $j$. This argument might seem trivial at first, but
the explicit consideration of this condition in the probabilistic
representation of ecological interactions will prove fundamental to understand
their variation. We define $X_{iy}$ as a stochastic variable representing the
occurrence of species $i$ at location $y$, and similarly $X_{ijy}$ the co-occurrence
of species $i$ and $j$. The quantity we seek to understand is the probability of a
joint event, conditional on the set of environmental conditions $E_y$:

%-----------------
\begin{equation}
	\text{P}(X_{i,y},X_{j,y},L_{ij,y},)
\end{equation}
%-----------------

Or simply said, the probability of observing both species $i$ and $j$, and an
interaction between $i$ and $j$. This probability could be decomposed in two parts
using the product rule of probabilities:

%-----------------
\begin{equation}
	P(X_{iy},X_{jy},L_{ijy})=P(X_{iy},X_{jy}|E_y)P(L_{ijy}|\mathbf{T},X_{iy},X_{jy},E_y)
\end{equation}
%-----------------

The left term is the probability of observing the two species co-occurring at
location $y$. It corresponds to the Grinnellian dimension of the niche. The
right term is a conditional probability, representing the probability that an
interaction occurs between species $i$ and $j$, given their set of traits $T$ and
that they are co-occurring. Above, we referred to this entity as the metaweb,
corresponding to the Eltonian dimension of the niche . Below, we will see how
this formalism can be directly fitted to empirical data. But before turning to
an application, we will discuss the interpretation of different variants of
these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses about temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{ijy} |E_y)$. In
this situation there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because of an impact
of ecological interactions on distributions \citep{Pollock2014} or, alternatively,
because of environmental requirements shared between i and j. In the former
case, species are not independent of each other and the conditional occurrence
must be accounted for explicitly, $P(X_{ijy} |E_y)=P(X_{iy} |E_y,X_{jy})P(X_{jy} |E_y)$.
In the latter case, species are independent and only the marginal occurrence
must be accounted for, $P(X_{ijy} |E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1976, Connor1980,
Gotelli2000}. Second- and higher-order interactions (e.g. trophic cascades)
could also affect co-occurrence. Co-occurrence of multiple species embedded in
ecological networks is however a topic of its own, influenced by the network
topology and species richness \citep{Cazelles2015}. Not only direct interactions do
influence co-occurrence, but indirect interactions as well (e.g. plant species
sharing an herbivore could repulse each other in space. The impact of direct
interactions and first-order indirect interactions on co-occurrence tends to
vanish with increasing species richness in the community. Further, co-
occurrence is also influenced by the covariance of interacting species to an
environmental gradient \citep{Cazelles2016}. Because of the complexity of relating
co-occurrence to the structure of interaction networks, we will here focus on
the variation of interactions and not on their distribution, and leave this
specific issue for the Discussion and future research.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, not
probabilistic (e.g. \citealt{Havens1992, Woods2015}). Species are assumed to interact
whenever they are found together in a location, independently of their
abundance and the environment. In other words, $P(L_{ijy}|X_{ijy}=1)$ equals 1
andit equals 0 if $P(L_{ijy}|X_{ijy}=0)$. This approach might be a reasonable
approximation if the scale of sampling and inference is so large that the
probability of observing at least one interaction converges to unity and the
only variation in the networks considered arises from species distributions.

Ecological interactions could also vary with the environment, so that $P(L_{ijy}
|E_y)$. Although it is not common to see a conditional representation of
pairwise ecological interactions, experimental studies have revealed them to
frequently be sensitive to the environment. For instance, \citep{Mckinnon2010} shown
that predation risks of shorebirds vary at the continental scale, decreasing
from the south to the north. It is also common to see increasing top-down
control with temperature (e.g. \citealt{Shurin2012, Gray2016}). Effects of the
environment on interactions propagate up the community and influence network
structure \citep{Woodward2010; Petchey2010}.

@@

We now turn to an illustration of the framework with the analysis of an
empirical dataset of host–parasite networks sampled throughout south–north
gradient in continental Europe. The focal system consists of local food webs
of willows (genus Salix), their galling insects, and the natural enemies of
these gallers. Targeting this system, we ask: i) how much does network
structure vary across the gradient, and ii) what is the primary driver of
network turnover across the gradient?

@@

Communities of willows, gallers, and galler enemies are species-rich and
widely distributed, with pronounced variation in community composition across
space. The genus Salix includes over 400 species, most of which are shrubs or
small trees \citep{Argus1997}. The genus is common in most habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}. Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The enemy community of the gallers includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2000}. These enemies encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent “extended phenotypes” of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumptively constrained by the form,
size, and thickness of its gall.

>>>>

% TODO:
% Add Ben and Spencer's edits
% Revise the model with fixed interaction pribabilites (havens)
% Compile the references
% Distribute for a last round of edits
% Add scussion on other on other types of interactions

@@

\textbf{Title:} Bringing Elton and Grinnell together: a quantitative framework to represent the biogeography of ecological interaction networks

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Ben Baiser$^{3}$, Jennifer A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo
Martinez$^{6}$, Tommy Nyman$^{7}$, Timoth\'ee Poisot$^{2,8}$,  Spencer A. Wood$^{9}$, Daniel B. Stouffer$^{10}$, Jason Tylianakis$^{10,11}$ Tomas Roslin$^{12}$,\\

@@

9: University of Washington, School of Environmental and Forest Sciences, 
Box 352100, Seattle, WA 98195, USA\\

@@

Biogeography has historically focused on the spatial distribution and
abundance of species, neglecting variation in the way species interact with
each other. Models of species distribution and of interactions have
historically taken different paths. Here, we plea for an integrated approach,
adopting the view that community structure is best represented as an
ecological network of interactions. We outline a conceptual approach,
suggesting that the ecological niche may be redefined to encompass the effect
of the environment on species distribution (the Grinnellian dimension of the
niche) and on the ecological interactions among them (the Eltonian dimension).
Starting from this novel concept, we develop a quantitative theory to explain
turnover of interactions in space and time – i.e. a novel approach to
interaction distribution modelling. We apply this novel framework to a large
data set of host–parasite interactions across Europe and find that two aspects
of the environment (temperature and precipitation) leave a strong imprint on
species co-occurrence, but not on the probability of local interactions. Even
where species co-occur, interaction proves a stochastic rather than
deterministic process, adding to variation in realized network structure. We
also find that a large majority of species pairs are never found together,
thus precluding any inferences regarding their probability to interact. Our
framework provides a first conceptual framework to explain the variation of
network structure at large spatial scales and opens new perspectives at the
frontier between species distribution modelling and community ecology.

@@

Community ecology is defined in most textbooks as \textit{the study of the
interactions that determine the distribution and abundance of organisms}
\citep{Krebs2001}. Despite a general consensus on this definition
(Scheiner2007), research on variation in community
structure has mostly focused on the spatial and temporal turnover of species composition
\citep{Anderson2011}, neglecting variation in the way species interact with
each other \citep{Poisot2015a}. Given this omission, it is perhaps not
surprising that biogeographers are still struggling to establish whether
interactions actually impact the distribution of species at large spatial
scales\citep{Wisz2012, Kissling2012}, even though there is enough evidence
from community ecology that interactions are critical for determining species
dynamics, resource use and responses to the environment af finer scales.
Recent attempts at accounting for interactions in species distribution models
\citep{Pollock2014, Pelissier2013} have brought some methodological advances.
Yet, these techniques are still based on a `species-based' approach to
communities, where interactions are merely treated as fixed covariates
affecting distribution.

As a more explicit description of interactions among species, the
network approach offers a convenient representation of communities. Species
are represented as nodes and interactions by links. To date, studies of
network diversity have mostly been concerned with the distribution of
interactions within locations, and less so with the variation among locations
\citep{Dunne2005, Bascompte2007, Ings2007, Kefi2012}. There is however 
ample evidence that interaction networks vary in space and time
\citep{Laliberté2010, Poisot2012, Albouy2014, Poisot2016, Trojelsgaard2015}. Metacommunity
theory provides explanation for the variation in the distribution of the
different nodes \citep{Gravel2011c, Pillai2011}, but there is no explanation
to the joint variation of node and link occurrences. We urgently need a
conceptual framework to formalize these observations.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is an evident need to bring the two together.
Here, we offer an integrated approach, adopting the view that community
structure is best represented as an ecological network of interactions. Based
on this idea, we propose a new description of the basic concept of the
ecological niche, now integrating the effect of the environment on species
distribution and on the ecological interactions among them. Starting from this
redefined concept, we develop a quantitative theory to explain turnover of
interactions in space and time. We first present the conceptual framework, and
then formalize it mathematically, using a probabilistic model to represent the
sampling of the regional pool of interactions. At the level of species pairs,
the statistical approach could be conceived as an interaction distribution
model. At the community level, the approach provides a likelihood-based method
to compare different hypotheses of network turnover. We apply this novel
framework to a large data set on host–parasite interactions across Europe and
find that variation of the environment causes turnover of both species and
interactions. The network structure changes systematically across the
latitudinal gradient, with a peak of connectance at intermediate latitude.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community
\citep{Gotzenberger2012}?} This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Figure 1. The metaweb represents
potential interactions among all species that could be found in a given area.
In this particular case, there are 275 nodes, and 1173 links among plants (52
nodes), herbivores (96 nodes), and parasitoids (127 nodes) from Northern
Europe. An instance of a local community is also illustrated, with 45 nodes
and 93 interactions. Only 55.0\% of all potential interactions are realized,
revealing the stochastic nature of ecological interactions. Our objective here
is to provide a conceptual framework to explain the sampling of the regional
pool of interactions, along with a quantitative method to predict it. The
problem could be formalized sequentially by first understanding why only a
fraction of the species are co-occurring locally and second why these species
do or do not interact.

There are multiple causes of spatial turnover in community composition. The
first and most-studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among sites by selecting the locally
fittest species \citep{Leibold2004}. Stochasticity plays an additional role,
either colonization and extinction events \citep{Hanski1999} are inherently
unpredictable, or because of strong non-linear feedbacks in community dynamics generating alternative
transients and equilibria \citep{Chase2007, Vellend2014}. Analyses of
community turnover are usually performed with data represented in a table with
rows corresponding to sites (or measurements) and columns to species. Metrics
of beta diversity quantify the variance of this community data
\citep{Legendre2005}. Traditional approaches rely on measures of dissimilarity
among communities, such as the Jaccard or Bray–Curtis indices. A more recent
approach decomposes the total variation of the community data into species and
site contributions to beta diversity \citep{Legendre2013}. Even though these
methods compare whole lists of species among sites or measurements, they
remain fundamentally ’species-based’, since they report the variation within
columns. None of them explicitly considers the variation of associations
(i.e., of pairs or higher-order motifs – \citealt{Stouffer2007}).

The "niche" is by far the dominant concept to explain species distributions
and community assembly, from the local to the global scale. Following
\citealt{Hutchinson1957}, the niche is viewed as the set of environmental
conditions allowing a population to establish and persist (see also
\citealt{Holt2009}). Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to operationalize with species distribution models, as it maps
naturally on available distributional and environmental data. In consequence,
a vast array of statistical tools have been developed to implement it (e.g.
BIOMOD \citealt{Thuiller2003}, MaxEnt \citealt{Phillips2006}). It is however
much harder to account for ecological interactions with this approach
\citep{Peterson2011}. As such, these interactions are often viewed as externalities
constraining or expanding the range of environmental conditions required for a
species to maintain a viable population \citep{Pulliam2000, Soberon2007, Boulangeat2012}.

@@

Interestingly, the ecological network literature also has its own ’niche
model’ to position a species in a community \citep{Williams2000}. The niche of a
species in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over 3 to 5 different niche axes \citep{Williams2000, Eklof2013}.
The niche model of food web structure and its variants have successfully explained the
complexity of a variety of networks, from food webs to plant–pollinator
systems \citep{Allesina2008, Williams2010, Eklof2013}. This
conceptual framework is, however, limited to local communities, and does not
provide any explanation for the turnover of network structure along
environmental gradients.

@@

A more integrative description of the niche is key to understand spatial and
temporal turnover in community structure. Despite several attempts to update
the concept of the ecological niche, ecologists have not moved far past the
``n-dimensional hypervolume'' defined by Hutchinson. Despite its intuitive
interpretation and easy translation into species distribution models
\citep{Boulangeat2012, Blonder2014}, the concept has been frequently
criticized \citep{Hardin1960, Peters1991, Silvertown2004}, and several
attempts have been made to expand and reinforce it \citep{Pulliam2000,
Chase2003, Soberon2007, Holt2009, McInerny2012b}.

Part of the problem surrounding the niche concept has been clarified with the
distinction between Eltonian and Grinnellian definitions \citep{Chase2003}. The
Grinnellian dimension of the niche is the set of environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is intuitive to apply, and constitutes the conceptual backbone of
species distribution models. The Eltonian niche, on the other hand, is the
effect of a species on its environment. This aspect of the niche is well known
by community ecologists, but is trickier to turn into predictive models.
Nonetheless, the development of the niche model of food web structure
\citep{Williams2000} and its parameterization \citep{Williams2010; Gravel2013}
made it more operational.

These perspectives are rather orthogonal to each other, which has resulted in
considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile with the following definition:
``[The niche is] the joint description of the environmental conditions that
allow a species to satisfy its minimum requirements so that the birth rate of
a local population is equal to or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions''.
Their representation merges zero-net growth isoclines delimiting the
Grinnellian niche (when does the population persists) with impact vectors delimiting
the Eltonian niche (what is the per-capita impact). While this representation
has been very influential in local-scale community ecology (the resource-ratio
theory of coexistence, \citealt{Tilman1982}), it remains impractical at larger
spatial scales because of the difficulties to measure it. The absence of any
mathematical representation of the niche that could easily be fitted to
ecological data may explain why biogeographers are still struggling to develop
species distribution models that also consider ecological interactions.

We propose to integrate the two perspectives of the niche with a visual
representation of both components (Fig. 2). The underlying rationale is that, in
addition to the environmental constraints on demographic performance, any
organism requires resources to sustain its metabolic demands and reproduction.
Abiotic environmental axes are any non-consumable factors affecting the
demographic performance of an organism. Alternatively, the resource axes are
traits of the resources that allow interactions with the consumer. The niche
should therefore be viewed as the set of abiotic environmental factors (the
Grinnellian component) along with the set of traits (the Eltonian component)
that allow a population to establish and to persist at a location. Accordingly,
each species can be characterized by an optimal position in both the
environmental (x-axis) and the trait (y-axis) plane. The integrated niche is
then the hypervolume where interactions can occur and sustain a population.
This approach radically changes the representation of the niche, putting the
distributions and ecological interactions of species in the same formalism.
The limits of the niche axes could be independent of each other (as in the
example of Fig. 2), or alternatively interact. For instance, the optimal
prey size for predatory fishes could decline with increasing temperature
\citep{Lelong2015}, which would make diet boundaries functions of the
environment. The other way around, we could also consider that the growth rate
of the predator changes with the size of its prey items, thereby altering the
environmental boundaries.

%========================================================%
\section*{A probabilistic representation of interaction networks in space}

We now formalize the integrated niche with a probabilistic approach to
interactions and distributions. We seek to represent the probability that an
interaction between species $i$ and $j$ occurs at location $y$. We define
$L_{ijy}$ as a stochastic variable, and are looking at the probability that
this event occurs, $P(L_{ijy})$. The occurrence of an interaction is dependent
on the co-occurrence of species $i$ and $j$. This argument might seem trivial
at first, but the explicit consideration of this condition in the
probabilistic representation of ecological interactions will prove fundamental
to understand their variation. We define $X_{iy}$ as a stochastic variable
representing the occurrence of species $i$ at location $y$. The quantity we seek to
understand is the probability of a joint event, conditional on the set of
environmental conditions $E_y$:

%-----------------
\begin{equation}
	\text{P}(X_{iy},X_{jy},L_{ijy}|E_y)
\end{equation}
%-----------------

Or simply said, the probability of observing both species $i$ and $j$ plus an
interaction between $i$ and $j$ given the conditions $E_y$. This probability
could be decomposed into two parts using the product rule of probabilities:

%-----------------
\begin{equation}
	P(X_{iy},X_{jy},L_{ijy}|E_y)=P(X_{iy},X_{jy}|E_y)P(L_{ijy}|\mathbf{T},X_{iy},X_{jy},E_y)
\end{equation}
%-----------------

The first term on the right-hand side of the equation is the probability of
observing the two species co-occurring at location $y$. It corresponds to the
Grinnellian dimension of the niche. The second term represents the probability
that an interaction occurs between species $i$ and $j$, given their set of
traits $T$ and that they are co-occurring. Above, we refer to this entity
as the "metaweb" and it corresponds to the Eltonian dimension of the niche. Below,
we will see how this formalism can be directly fitted to empirical data. But
before turning to an application, we will discuss the interpretation of
different variants of these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses about temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{iy},X_{jy}|E_y)$.
In this situation there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because interactions
do constraint distribution \citep{Pollock2014, Cazelles2016} or,
alternatively, because of environmental requirements shared between $i$ and
$j$. In the former case, species are not independent of each other and the
conditional occurrence must be accounted for explicitly, $P(X_{iy},X_{jy}
|E_y)=P(X_{iy}|E_y,X_{jy})P(X_{jy}|E_y)$. In the latter case, species are
independent and only the marginal occurrence must be accounted for, $P(
|E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1976, Connor1980,
Gotelli2000}. Second- and higher-order interactions (e.g. trophic cascades)
could also affect co-occurrence. Co-occurrence of multiple species embedded in
ecological networks is a topic of its own, however, and is influenced by both network
topology and species richness \citep{Cazelles2016}. Not only direct interactions do
influence co-occurrence, but indirect interactions do as well (e.g. plant species
sharing an herbivore could repulse each other in space). The impact of direct
interactions and first-order indirect interactions on co-occurrence tends to
vanish with increasing species richness in the community. Further, co-occurrence
is also influenced by the covariance of interacting species to an
environmental gradient \citep{Cazelles2015}. Because of the complexity of relating
co-occurrence to the structure of interaction networks, we will here focus on
the variation of interactions and not on their distribution, and leave this
specific issue for the Perspectives and future research.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, not
probabilistic (e.g. \citealt{Havens1992, Wood2015}). Species are assumed to interact
whenever they are found together in a location, independent of their
local abundance and the local environment. In other words, $P(L_{ijy}|X_{ijy}=1) = 1$
and $P(L_{ijy}|X_{ijy}=0) = 0$. This approach might be a reasonable
approximation if the scale of sampling and inference is so large that the
probability of observing at least one interaction converges to unity and the
only variation in the networks considered arises from species distributions.

Ecological interactions could also vary with the environment, so that $P(L_{ijy}
|E_y)$. Although it is not common to see a conditional representation of
pairwise ecological interactions, experimental studies have revealed them to
frequently be sensitive to the environment. For instance, \citep{Mckinnon2010} showed
that predation risks of shorebirds vary at the continental scale, decreasing
from the south to the north. It is also common to see increasing top-down
control with temperature (e.g. \citealt{Shurin2012, Gray2016}). Effects of the
environment on interactions also propagate up the community and influence network
structure \citep{Tylianakis2007, Woodward2010; Petchey2010; Lelong2015}.

@@

We now turn to an illustration of the framework with the analysis of an
empirical dataset of host–parasite networks sampled throughout the south–north
environmental gradient in continental Europe. The focal system consists of local food webs
of willows (genus Salix), their galling insects, and the parasitoids of
these gallers. Targeting this system, we ask: i) how much does network
structure vary across the gradient, and ii) what is the primary driver of
network turnover across the gradient?

@@

Communities of willows, gallers, and parasitoids are species-rich and
widely distributed, with pronounced variation in community composition across
space. The genus Salix includes over 400 species, most of which are shrubs or
small trees \citep{Argus1997}. The genus is common in most habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}). Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The parasitoid community includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2000}. These encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent ``extended phenotypes'' of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumably constrained by the form,
size, and thickness of its gall.
